\renewcommand*{\standardstate}{{-\kern-6pt{\ominus}\kern-6pt-}}

# Chemical Equilibrium {#ChemicalEquilibrium}

In this chapter, we will concentrate on chemical processes that happen at constant $T$ and constant $P$.^[The majority of chemical reactions in a lab happens at those conditions, and all biological functions happen at those conditions as well.] As such, we will focus our attention on the Gibbs free energy.

## Gibbs Equation {#gibbseqsec}
Recalling from the previous chapter, the definition of $G$ is:

\begin{equation}
G = U -TS -PV = H-TS,
(\#eq:dgdeftot)
\end{equation}

which, taking the differential at constant $T$ and $P$, becomes:

\begin{equation}
\begin{aligned}
dG &= dU \; \overbrace{-SdT}^{=0} -TdS \; \overbrace{-VdP}^{=0} -PdV = dH \; \overbrace{-SdT}^{=0} -TdS \\
   &= dH -TdS.
\end{aligned}
(\#eq:dgdeftot2)
\end{equation}

Integrating eq. (\#eq:dgdeftot2) between the initial and final states of a process results in:

\begin{equation}
\begin{aligned}
\int_i^f dG &= \int_i^f dH -T \int_i^f dS \\
\\
\Delta G &= \Delta H -T \Delta S
\end{aligned}
(\#eq:gibbseq)
\end{equation}


which is the famous **Gibbs equation** for $\Delta G$. Using Definition \@ref(def:helmgibbsminimum), we can use $\Delta G$ to infer the spontaneity of a chemical process that happens at constant $T$ and $P$ using $\Delta G \leq 0$. If we set ourselves at standard conditions, we can calculate the standard Gibbs free energy of formation, $\Delta_{\text{rxn}} G^\standardstate$, for any reaction as:

\begin{equation}
\begin{aligned}
\Delta_{\text{rxn}} G^\standardstate &= \Delta_{\text{rxn}} H^\standardstate -T \Delta_{\text{rxn}} S^\standardstate \\
\\
&= \sum_i \nu_i \Delta_{\mathrm{f}} H_i^\standardstate + T \sum_i \nu_i S_i^\standardstate,
\end{aligned}
(\#eq:gibbseq2)
\end{equation}

where $\Delta_{\mathrm{f}} H_i^\standardstate$ are the standard enthalpies of formation, $S_i^\standardstate$ are the standard entropies, and $\nu_i$ are the stoichiometric coefficients for every species $i$ involved in the reaction. All these quantities are commonly available, and we have already discussed their usage in Chapters \@ref(Thermochemistry) and \@ref(thirdlaw), respectively.^[It is not uncommon to see values of $\Delta_{\text{f}} G^\standardstate$ tabulated alongside $\Delta_{\mathrm{f}} H^\standardstate$ and $S_i^\standardstate$, which simplifies even further the calculation. If $\Delta_{\text{f}} G^\standardstate$ are not reported, they can always be calculated by their constituents.]

The following four possibilities are possible for $\Delta G^\standardstate$ of a chemical reaction:

| $\Delta G^\standardstate$ |      | $\Delta H^\standardstate$ | $\Delta S^\standardstate$ |                             Spontaneous?                             |
|:-------------------------:|------|:-------------------------:|:-------------------------:|:--------------------------------------------------------------------:|
|             –             | *if* |             –             |             +             |                               *Always*                               |
|             +             | *if* |             +             |             –             |                                *Never*                               |
|            –/+            | *if* |             –             |             –             | *Depends on $T$:*  <br> $\scriptstyle{\text{spontaneous at low } T}$ |
|            +/–            | *if* |             +             |             +             | *Depends on $T$:* <br> $\scriptstyle{\text{spontaneous at high } T}$ |

Or, in other words:

- Exothermic reactions that increase the entropy are always spontaneous.
- Endothermic reactions that reduce the entropy are always non-spontaneous.
- For the other two cases, the spontaneity of the reaction depends on the temperature:
    - Exothermic reactions that reduce the entropy are spontaneous at low $T$. 
    - Endothermic reactions that increase the entropy are spontaneous at high $T$.

    
A simple criterion to evaluate the entropic contribution of a reaction is to look at the total number of moles of the reactants and the products (as the sum of the stoichiometric coefficients). If the reaction is producing more molecules than it destroys $\left( \left| \sum_\text{products} \nu_i \right| > \left| \sum_\text{reactants} \nu_i \right| \right)$, it will increase the entropy. Vice versa, if the total number of moles in a reaction is reducing $\left( \left| \sum_\text{products} \nu_i \right| < \left| \sum_\text{reactants} \nu_i \right| \right)$, the entropy will also reduce.


XXX exercise to calculate T inversion XXX


## Behavior of the Gibbs Free Energy {#Gbehavior}
As we saw in Section \@ref(thermpot), the natural variables of the Gibbs free energy are the temperature, $T$, the pressure, $P$, and chemical composition, as the number of moles $\{n_i\}$. The Gibbs free energy can therefore be expressed using the total differential as (see also, last formula in eq. \@ref(eq:dhagchem2)):

\begin{equation}
dG(T,P,\{n_i\}) = \mkern-18mu \underbrace{\left(\frac{\partial G}{\partial T} \right)_{V,\{n_i\}}}_{\text{temperature dependence}} \mkern-36mu dT + \underbrace{\left(\frac{\partial G}{\partial P} \right)_{T,\{n_i\}}}_{\text{pressure dependence}} \mkern-36mu dP + \sum_i \underbrace{\left(\frac{\partial G}{\partial n_i} \right)_{T,P,\{n_{j \neq i}\}}}_{\text{composition dependence}} \mkern-36mu dn_i\;.
(\#eq:dgtotdiff)
\end{equation}

If we know the behavior of $G$ as we vary each of the three natural variables independently of the other two, we can reconstruct the total differential $dG$. Each of these terms represents a coefficient in eq. \@ref(eq:dgtotdiff), which are given in eq. \@ref(eq:dhagchem3), and reported and discussed in the sections below.

### Temperature dependence of $\Delta G$
$$
\left(\frac{\partial G}{\partial T} \right)_{P,\{n_i\}}=-S
$$

Let's analyze the first coefficient that gives the dependence of the Gibbs energy on temperature.  Since this coefficient is equal to $-S$ and the entropy is always positive, $G$ must decrease when $T$ increases at constant $T$ and $\{n_i\}$, and vice versa.

If we replace this coefficient for $-S$ in the Gibbs equation (eq. \@ref(eq:gibbseq)), we obtain:

\begin{equation}
\Delta G = \Delta H + T \left(\frac{\partial G}{\partial T} \right)_{P,\{n_i\}},
(\#eq:dgconsttt)
\end{equation}

and since eq. \@ref(eq:dgconsttt) includes both $\Delta G$ and its partial derivative with respect to temperature $\left(\frac{\partial \Delta G}{\partial T} \right)_{P,\{n_i\}}$ we need to rearrange it to include the temperature derivative only. To do so, we can start by evaluating the partial derivative of $\frac{\Delta G}{T}$ using the chain rule:

\begin{equation}
\left[ \frac{\partial\left( \frac{\Delta G}{T} \right)}{\partial T} \right]_{P,\{n_i\}} = \frac{1}{T} \left(\frac{\partial \Delta G}{\partial T}  \right)_{P,\{n_i\}} - \frac{1}{T^2}\Delta G,
(\#eq:dgconsttt1)
\end{equation}

which, replacing $\Delta G$ from eq. \@ref(eq:dgconsttt) into eq. \@ref(eq:dgconsttt1), becomes:

\begin{equation}
\begin{aligned}
\left[ \frac{\partial\left( \frac{\Delta G}{T} \right)}{\partial T} \right]_{P,\{n_i\}} &= \frac{1}{T} \left(\frac{\partial \Delta G}{\partial T}  \right)_{P,\{n_i\}} - \frac{1}{T^2} \left[  \frac{1}{T} \left(\frac{\partial \Delta G}{\partial T}  \right)_{P,\{n_i\}} - \frac{1}{T^2}\Delta G \right] \\
&= \frac{1}{T} \left(\frac{\partial \Delta G}{\partial T}  \right)_{P,\{n_i\}}-\frac{1}{T} \left(\frac{\partial \Delta G}{\partial T}  \right)_{P,\{n_i\}} - \frac{\Delta H}{T^2},
\end{aligned}
(\#eq:dgconsttt2)
\end{equation}

which simplifies to:

\begin{equation}
\begin{aligned}
\left[ \frac{\partial\left( \frac{\Delta G}{T} \right)}{\partial T} \right]_{P,\{n_i\}} &= - \frac{\Delta H}{T^2}.
\end{aligned}
(\#eq:gibbshelm)
\end{equation}

Equation \@ref(eq:gibbshelm) is known as the **Gibbs–Helmholtz equation**, and is useful in its integrated form to calculate the Gibbs free energy for a chemical reaction at any temperature $T$ by knowing just the standard Gibbs free energy of formation and the standard enthalpy of formation for the individual species, which are usually reported at $T=298\;\text{K}$:

\begin{equation}
\begin{aligned}
\int_{T_i=298 \;\text{K}}^{T_f=T}  \frac{\partial\left( \frac{\Delta_{\text{rxn}} G}{T} \right)}{\partial T} &=  - \int_{T_i=298 \;\text{K}}^{T_f=T}  \frac{\Delta_{\text{rxn}} H}{T^2} \\ \\
\frac{\Delta_{\text{rxn}} G^\standardstate(T)}{T} &=  \frac{\Delta_{\text{rxn}} G^\standardstate}{298 \;\text{K}} + \Delta_{\text{rxn}} H^\standardstate \left( \frac{1}{T^2} -\frac{1}{(298 \;\text{K})^2} \right) \\
\\
\frac{\Delta_{\text{rxn}} G^\standardstate(T)}{T} &=  \frac{\sum_i \nu_i \Delta_{\text{f}} G_i^\standardstate}{298 \;\text{K}} + \sum_i \nu_i \Delta_{\text{f}} H_i^\standardstate \left( \frac{1}{T^2} -\frac{1}{(298 \;\text{K})^2} \right)
\end{aligned}
(\#eq:gibbshelmI)
\end{equation}


### Pressure dependence of $\Delta G$
$$
\left(\frac{\partial G}{\partial P} \right)_{T,\{n_i\}}=V
$$

We can now turn the attention to the second coefficient that gives how the Gibbs free energy changes when the pressure change. To do this, we put the system at constant $T$ and $\{n_i\}$, and then we consider infinitesimal variations of $G$. From eq. \@ref(eq:dgdef2):

\begin{equation}
dG = VdP -SdT +\sum_i\mu_i dn_i \quad \xrightarrow{\text{constant}\; T,\{n_i\}} \quad dG = VdP,
(\#eq:dgconsttp)
\end{equation}

which is the differential equation that we were looking for. To study changes of $G$ for macroscopic changes in $P$, we can integrate eq. \@ref(eq:dgconsttp) between initial and final pressures, and considering an ideal gas, we obtain:

\begin{equation}
\begin{aligned}
\int_i^f dG &= \int_i^f VdP \\
\Delta G &= nRT \int_i^f \frac{dP}{P} = nRT \ln \frac{P_f}{P_i}.
\end{aligned}
(\#eq:dgconsttpintegral1)
\end{equation}

If we take $P_i = P^\standardstate = 1 \, \text{bar}$, we can rewrite eq. \@ref(eq:dgconsttp) as:

\begin{equation}
G = G^\standardstate + nRT \ln \frac{P_f}{P^\standardstate},
(\#eq:dgconsttpintegral2)
\end{equation}

which is useful to convert standard Gibbs free energies of formation at pressures different than standard pressure, using:

\begin{equation}
\Delta_{\text{f}} G = \Delta_{\text{f}} G^\standardstate + nRT \ln \frac{P_f}{\underbrace{P^\standardstate}_{=1 \; \text{bar}}} = \Delta_{\text{f}} G^\standardstate + nRT \ln P_f
(\#eq:dgconsttpintegral2b)
\end{equation}


For liquids and solids, $V$ is essentially independent of $P$ (liquids and solids are incompressible), and eq. \@ref(eq:dgconsttp) can be integrated as:

\begin{equation}
\Delta G = \int_i^f VdP = V \int_i^f dP = V \Delta P.
(\#eq:dgconsttpintegral3)
\end{equation}

The plots in Figure \@ref(fig:gdeppfig) show the remarkable difference in the behaviors of $\Delta_{\text{f}} G$ for a gas and for a liquid, as obtained from eqs. \@ref(eq:dgconsttpintegral1) and \@ref(eq:dgconsttpintegral3).

```{r gdeppfig, echo=FALSE, fig.align='center', fig.cap='Dependence of the Gibbs Free Energiy of Formation of Liquid and Gaseous Ethanol at T=310 K. The Curves Cross at the Vapor Pressure of Liquid Ethanol at T=310 K, which is 0.1 bar'}
x <- seq(0.01,10,0.1)
i <- 0.2
f <- 1
p1 <- 0.5
plot(x, -168+8.31*10^(-3)*298*log(x),
     xlim=c(0.3, 9.7),
     ylim=c(-180, -160),
     type="l",
     ylab = expression(paste(Delta, "G [kJ/mol]")), xlab = "Pressure [bar]", col = "#2E9FDF")
arrows(-10,-174, 20, -174 , length=0, angle=90, code=3)
text(8.07,-164, "gas", col = "#2E9FDF")
text(8.07,-173, "liquid", col="black")
points(0.1,-174, col="red", pch=19)
text(1.16,-173, expression(paste('P'[vap], "* = 0.1 bar")), col = "red")
legend("topleft", inset=.05, expression(paste("1 mol of ",C[2],H[5],OH, " at constant T = 310 K")),box.col = "white", bg = "white", adj=0.15)
```

XXX exercise on P vap xxx

### Composition dependence of $\Delta G$
$$
\left(\frac{\partial G}{\partial n_i} \right)_{T,P}=\mu_i
$$

The third and final coefficient gives the chemical potential as the dependence of $G$ on the chemical composition at constant $T$ and $P$. Similarly to the previous cases, we can take the definition of the coefficient and integrate it directly between the initial and final stages of a reaction. If we consider a reaction product, pure substance $i$, at the beginning of the reaction there will be no moles of it $n_i=0$, and consequently $G=0$.^[For reactants, the same situation usually applies but in reverse. More complicated cases where the reaction does not consume all reactants are possible, but insignificant for the following treatment.] For convenience, we can then term the number of moles of product $i$ at the end of the reaction $n$, and its Gibbs free energy $G$. The integral will become:

\begin{equation}
\int_0^G d G = \int_0^n \mu^* dn,
(\#eq:dgconstn)
\end{equation}

where $\mu^*$ indicates the chemical potential of a pure substance, which is independent on the number of moles by definition. As such, eq. \@ref(eq:dgconstn) becomes:

\begin{equation}
\int_0^G d G = \mu^* \int_0^n dn \quad \rightarrow \quad G = \mu^* n \quad \rightarrow \quad \mu^* = \frac{G}{n},
(\#eq:dgconstn1)
\end{equation}

which gives a simple interpretation of the chemical potential of a pure substance as the molar Gibbs free energy.

We can start from eq. \@ref(eq:dgconsttpintegral2) and write for a pure substance that is brought from $P_i=P^\standardstate$ to $P_f=P$ at constant $T$:

\begin{equation}
G - G^\standardstate = nRT \ln \frac{P}{P^\standardstate},
(\#eq:dgconstn2)
\end{equation}

dividing both sides by $n$, we obtain:

\begin{equation}
\frac{G}{n} - \frac{G^\standardstate}{n} = RT \ln \frac{P}{P^\standardstate},
(\#eq:dgconstn3)
\end{equation}

which, for a pure substance at $P^\standardstate = 1 \;\text{bar}$, becomes:

\begin{equation}
\mu^* = \mu^\standardstate + RT \ln P.
(\#eq:dgconstn4)
\end{equation}

Let's now consider a mixture of ideal gases, and let's try to find out whether the chemical potential of a pure gas inside the mixture, $\mu_i^{\text{mixture}}$, is the same as its chemical potential outside the mixture, $\mu^*$.
To do so, we can use eq. \@ref(eq:dgconstn4) and replace the pressure $P$ with the partial pressure  $P_i$:

\begin{equation}
\mu_i^{\text{mixture}} = \mu_i^\standardstate + RT \ln P_i,
(\#eq:dgconstn5)
\end{equation}

where the partial pressure $P_i$ can be obtained from the simple relation that is known as **Dalton's Law**:

\begin{equation}
P_i = y_i P,
(\#eq:daltonlaw)
\end{equation}

with $y_i$ being the concentration of gas $i$ measured as a mole fraction in the gas phase $y_i=\frac{n_i}{n_{\text{TOT}}} < 1$. Replacing eq. \@ref(eq:daltonlaw) into eq. \@ref(eq:dgconstn5), we obtain:

\begin{equation}
\begin{aligned}
\mu_i^{\text{mixture}} &= \mu_i^\standardstate + RT \ln (y_i P) \\
 &= \underbrace{\mu_i^\standardstate + RT \ln P}{\mu_i^*} + RT \ln y_i,
\end{aligned}
(\#eq:dgconstn6)
\end{equation}

which then reduces to the following equation:

\begin{equation}
\mu_i^{\text{mixture}} = \mu_i^* + RT \ln y_i.
(\#eq:dgconstn7)
\end{equation}

Analyzing eq. \@ref(eq:dgconstn7), we can immediately see that, since $x_1 < 1$:

\begin{equation}
\mu_i^{\text{mixture}} < \mu_i^*,
(\#eq:dgconstn8)
\end{equation}

or, in other words, the chemical potential of a substance in the mixture will be lower than the chemical potential of the pure substance. If we consider a process where we start from two separate pure ideal gases, and we finish it with a mixture of the two, we can calculate the change in Gibbs free energy due to the mixing process as:

\begin{equation}
\Delta_{\text{mixing}} G = \sum n_i \left( \mu_i^{\text{mixture}} - \mu_i^* \right)  < 0,
(\#eq:dgconstn7)
\end{equation}

or, in other words, the process is spontaneous under all circumstances, and pure ideal gases will always mix.

## Chemical Equilibrium {#chemeq}

Let's consider a prototypical reaction at constant $T,P$:

\begin{equation}
 a\mathrm{A} + b\mathrm{B} \rightarrow c\mathrm{C} + d\mathrm{D} 
(\#eq:reaction4)
\end{equation}

The Gibbs free energy of the reaction is defined as:

\begin{equation}
\Delta_{\text{rxn}} G = G_{\text{products}} - G_{\text{reactants}} = G^{\text{C}} + G^{\text{D}} - G^{\text{A}}-G^{\text{B}},
(\#eq:chemeq1)
\end{equation}

and replacing the absolute Gibbs free energies with the chemical potentials $\mu_i$, we obtain:

\begin{equation}
\Delta_{\text{rxn}} G = c \mu_{\text{C}} + d \mu_{\text{D}} - a \mu_{\text{A}}- b\mu_{\text{B}}.
(\#eq:chemeq2)
\end{equation}

Assuming the reaction is happening in the gas phase, we can then use eq. \@ref(eq:dgconstn5) to replace the chemical potentials with their value in the reaction mixture, as:

\begin{equation}
\begin{aligned}
\mkern-60mu \Delta_{\text{rxn}} G =& \; c (\mu_{\text{C}}^\standardstate +RT \ln P_{\text{C}}) + d (\mu_{\text{D}}^\standardstate +RT \ln P_{\text{D}}) +\\ & - a (\mu_{\text{A}}^\standardstate +RT \ln P_{\text{A}}) - b (\mu_{\text{B}}^\standardstate +RT \ln P_{\text{B}}) \\
  =& \; \underbrace{c \mu_{\text{C}}^\standardstate + d \mu_{\text{D}}^\standardstate - a \mu_{\text{A}}^\standardstate - b\mu_{\text{B}}^\standardstate}_{\Delta_{\text{rxn}} G^\standardstate}  +RT \ln \frac{P_{\text{C}}^c \cdot P_{\text{D}}^d}{P_{\text{A}}^a \cdot P_{\text{B}}^b}.
\end{aligned}
(\#eq:chemeq3)
\end{equation}

If we define a new quantity called the *reaction quotient* as a function of the partial pressures of each substance:

\begin{equation}
Q_P = \frac{P_{\text{C}}^c \cdot P_{\text{D}}^d}{P_{\text{A}}^a \cdot P_{\text{B}}^b},
(\#eq:chemeq4)
\end{equation}

we can then simply rewrite eq. \@ref(eq:chemeq3) using eq. \@ref(eq:chemeq4) as:

\begin{equation}
\Delta_{\text{rxn}} G = \Delta_{\text{rxn}} G^\standardstate + RT \ln Q_P.
(\#eq:chemeq5)
\end{equation}

This equation tells us that the sign of $\Delta_{\text{rxn}} G$ is influenced by the reaction quotient $Q_P$. For a spontaneous reaction at the beginning, the partial pressures of the reactants are much higher than the partial pressures of the porducts, therefore $Q_P \ll 1$ and $\Delta_{\text{rxn}} G < 0$, as we expect. As the reaction proceeds, the partial pressures of the products will increase, while the partial pressures of the reactants will decrease. Consequently, both $Q_P$ and $\Delta_{\text{rxn}} G$ will increase. The reaction will completely stop when $\Delta_{\text{rxn}} G = 0$, which is the chemical equilibrium point. At the reaction equilibrium:

\begin{equation}
\Delta_{\text{rxn}} G = 0 = \Delta_{\text{rxn}} G^\standardstate + RT \ln K_P,
(\#eq:chemeq6)
\end{equation}

where we have defined a new quantity called reaction constant, as the value the reaction quotient assumes when the reaction reaches equilibrium, and we have denoted it with the symbol $K_P$.^[Notice that the subscript $P$ refers to the fact that the equilibrium constant is measured in terms of partial pressures.] From eq. \@ref(eq:chemeq6) we can derive the following fundamental equation on the standard Gibbs free energy of reaction:

\begin{equation}
\Delta_{\text{rxn}} G^\standardstate = - RT \ln K_P.
(\#eq:chemeqdg0)
\end{equation}

To extend the concept of $K_P$ beyond the four species in the prototypical reaction \@ref(eq:reaction4), we can use the product of a series symbol $\left( \prod_i \right)$, and write:

\begin{equation}
K_P=\prod_i P_{i,\text{eq}}^{\nu_i},
(\#eq:kpdef)
\end{equation}

where $P_{i,\text{eq}}$ are the partial pressure of each species at equilibrium. Eq. \@ref(eq:kpdef) is in principle valid for ideal gases only. However, reaction involving ideal gases are pretty rare. As such, we can further extend the concept of equilibrium constant and write:

\begin{equation}
K_{\text{eq}} =\prod_i a_{i,\text{eq}}^{\nu_i},
(\#eq:kdef)
\end{equation}

where we have replaced the partial pressure at equilibrium, $P_{i,\text{eq}}$, with a new concept introduced initially by Gilbert Newton Lewis (1875–1946),^[Gilber Lewis is the same scientist that invented the concept of Lewis Structures.] that he termed **activiy**, and represented by the letter $a$. For ideal gases, it is clear that $a_i=P_i$. For real gases, the activity is equal to the fugacity $a_i=f_i$, a concept that we will investigate in the next chapter. For pure liquids and solids, the activity is simply $a_i=1$. For diluted solutions, the activity is equal to a measured concentration (such as, for example, the mole fraction $x_i$ in the liquid phase, and $y_i$ in the gas phase, molality $m_i$, or molar concentration $[i]$). Finally for concentrated solutions, the activity is related to the measured concentration via an activity coefficient. We will return to the concept of activity in Chapter \@ref(Solutions), when we will specifically deal with solutions. For now, it is interesting to use the activity to write the definition of the following two constants:

\begin{equation}
K_y =\prod_i y_{i,\text{eq}}^{\nu_i} \qquad \qquad \qquad \qquad K_C =\prod_i [i]_{\text{eq}}^{\nu_i} \; ,
(\#eq:kxkcdef)
\end{equation}

which can then be related with $K_P$ for a mixture of ideal gases using:

\begin{equation}
P_i = y_i P \qquad \qquad \qquad P_i=\frac{n_i}{V}RT=[i]RT,
(\#eq:kxkcdef2)
\end{equation}

which then results in:

\begin{equation}
K_p = K_y\cdot P^{\nu} \qquad \qquad K_C = K_P \left( \frac{1}{RT} \right)^{\nu},
(\#eq:kxkcdef3)
\end{equation}

with $\nu =\sum_i \nu_i$. 

Using the general equilibrium constant, $K_{\text{eq}}$, we can also rewrite the fundamental equation on $\Delta_{\text{rxn}} G^\standardstate$ that we derived in eq. \@ref(eq:chemeqdg0) to be applicable at most conditions, as:

\begin{equation}
\Delta_{\text{rxn}} G^\standardstate = - RT \ln K_{\text{eq}},
(\#eq:chemeqdg0)
\end{equation}

and since $\Delta_{\text{rxn}} G^\standardstate$ depends on $T,P$ and $\{n_i\}$, it is useful to explore how $K_{\text{eq}}$ depends on those variables as well.


xxx eventually exercise xxxx

### Temperature dependence of $K_{\text{eq}}$

To study the temperature dependence of $K_{\text{eq}}$ we can use eq. \@ref(eq:chemeqdg0) for the general equilibrium constant and write:

\begin{equation}
\ln K_{\text{eq}} = -\frac{\Delta G^\standardstate}{RT},
(\#eq:kptemp1)
\end{equation}

which we can then differentiate with respect to $dT$ at constant $P,\{n_i\}$ on both sides:

\begin{equation}
\left( \frac{\partial \ln K_{\text{eq}}}{\partial T} \right)_{P,\{n_i\}} = -\frac{1}{R} \left[ \partial \left( \frac{\partial \frac{\Delta F^\standardstate}{RT}}{\partial T} \right) \right]_{P,\{n_i\}},
(\#eq:kptemp2)
\end{equation}

and, using Gibbs-Helmholtz equation (eq. \@ref(eq:gibbshelm)) to simplify the left hand side, becomes:

\begin{equation}
\left( \frac{\partial \ln K_{\text{eq}}}{\partial T} \right)_{P,\{n_i\}} = -\frac{1}{R} \left( -\frac{\Delta H^\standardstate}{T^2} \right) = \frac{\Delta H^\standardstate}{RT^2},
(\#eq:kptemp3)
\end{equation}

which gives the dependence of $\ln K_{\text{eq}}$ on $T$ that we were looking for. Eq. \@ref(eq:kptemp3) is also called **van 't Hoff equation**,^[named after Jacobus Henricus "Henry" van 't Hoff Jr. (1852–1911).] and it is the mathematical expression of Le Chatelier's principle. The simplest interpretation is as follows:

- For an exothermic reaction ($\Delta H^\standardstate < 0$): $K_{\text{eq}}$ will decrease as the temperature increases. 
- For an endothermic reaction ($\Delta H^\standardstate > 0$): $K_{\text{eq}}$ will increase as the temperature increases. 

If we integrate the van 't Hoff equation between two arbitrary points at constant $P$, and assuming constant $\Delta H^\standardstate$, we obtain the following:

\begin{equation}
\int_1^2 d \ln K_{\text{eq}} = \frac{\Delta H^\standardstate}{R} \int_1^2 \frac{dT}{T^2},
(\#eq:kptemp4)
\end{equation}

which leads to the linear equation:

\begin{equation}
\ln K_{\text{eq}}(2) = \ln K_{\text{eq}}(1) - \frac{\Delta H^\standardstate}{R} \left( \frac{1}{T_2}-\frac{1}{T_2} \right).
(\#eq:vthoffeqI)
\end{equation}

which is the equation that produces the so-called **van 't Hoff plots**, from which $\Delta H^\standardstate$ can be experimentally determined:

```{r vthoffplt, echo=FALSE, fig.align='center', fig.cap='Van \'t Hoff Plots for an Endothermic (Reft, Blue) and an Exothermic (Right, Red) Reactions at Constant P.'}
x <- seq(0.2,0.8,0.01)
i <- 0.2
f <- 1
p1 <- 0.5
old.par <- par(mfrow=c(1, 2))
par(pty="s")
par(mar=c(0,2,0,2))
plot(x, 90-90*x,
     xlim=c(0.2, 0.8),
     ylim=c(0, 100),
     type="l",
     xaxt='n',
     yaxt='n',
     mgp=c(0.5,0.5,0),
     ylab = "ln K", xlab = "1/T", col = "#2E9FDF")
arrows(0.3 , 35, 0.65, 35 , length=0, angle=90, code=3, lty="dotted")
points(0.25,90-90*0.25, col="black", pch=19)
points(0.75,90-90*0.75, col="black", pch=19)
text(0.28,90-90*0.25+5, "1", col="black")
text(0.78,90-90*0.75+5, "2", col="black")
text(0.4,42, expression(paste(-Delta, H^o,"/T")), col = "black")
legend("topright", inset=.02, "Endothermic",, bg = "white", adj=0.15, box.col = "white")

par(mar=c(0,2,0,2))
plot(x, 10+90*x,
     xlim=c(0.2, 0.8),
     ylim=c(0, 100),
     type="l",
     xaxt='n',
     yaxt='n',
     mgp=c(0.5,0.5,0),
     ylab = "ln K", xlab = "1/T", col = "red")
arrows(0.3 , 40, 0.65, 40 , length=0, angle=90, code=3, lty="dotted")
points(0.25,10+90*0.25, col="black", pch=19)
points(0.75,10+90*0.75, col="black", pch=19)
text(0.23,10+90*0.25+5, "1", col="black")
text(0.73,10+90*0.75+5, "2", col="black")
text(0.5,47, expression(paste(-Delta, H^o,"/T")), col = "black")
legend("topright", inset=.02, "Exothermic",, bg = "white", adj=0.15, box.col = "white")
par(old.par)
```


### Pressure and composition dependence of $K_{\text{eq}}$

The two main conditions under which 


